---
title: "Religiosity as Influenced by Intersectionality"
shorttitle: "Paper"
abstract: An Abstract
keywords: [keyword1, keyword2]
thanks: Thanks to everyone for checking this out.
reference-section-title: References
bibliography: ../bibliography/project.bib
csl: ../bibliography/chicago-parenthetical.csl
format:
  aog-article-pdf:
    keep-tex: false
    include-in-header: 
      text: |
        \usepackage{dcolumn}
#  submittable-pdf:
#     keep-tex: false
#     fig-pos: "!t"
#     include-in-header: 
#       text: |
#         \usepackage{dcolumn}
  submittable-docx: default
---

```{r}
#| label: setup
#| include: false
library(here)
source(here("utils","check_packages.R"))
source(here("utils","functions.R"))
load(here("data","data_constructed","analytical_data.RData"))
```

# Introduction
Religious beliefs are an important component of social construction. The United States provides an interesting area to study trends in religious beliefs. The religious freedom in the United States has brought together a wide range of religious institutions, beliefs, and practices including no belief at all. However, these beliefs are not created and do not exist in a vacuum. Each component of society interacts with religious belief and vice-versa. For example, education, gender, race, socioeconomic status, and age are all factors both influencing religious belief and being influenced by religious belief and practices. The opportunity of religious freedom allows for each aspect of the human experience to influence and be influenced by trends in religiosity. 

Intersectionality is referred to in sociology as the interconnected nature of social categorizations such gender, race, and class creating overlapping and independent systems of advantage or disadvantage. Religiosity is a term describing the importance of and influence religious beliefs and practices have on individuals and society. We have seen in the United States changes in trends in religiosity over time. These changes are likely influenced by the interconnecting aspects of human identity. Previous studies have shown negative associations between levels of education and religiosity over time. Some studies have shown age having a positive influence on religiosity over time as well. Other studies have shown international results highlighting a negative association between religious importance and GDP as a socioeconomic factor, but found education to be a mediator. Some research has shown women to be more religious than men, on average, when considering intrinsic measure of religious importance and daily spiritual experiences. Other research has even shown interesting similatrities between the religiously affiliated and unaffiliated in the United States.

There are measurable trends in religious and spiritual beliefs and practices over the last century and they give rise to some interesting questions. What societal factors influence the change in religious and spiritual beliefs? How does this potentially change with age and between generations? How do these aspects work together to influence a persons measurable religiosity?  
  
# Background
One of the reasons much research has been done in this area is the social meaning of religious and spiritual beliefs are important and change over time. The social meaning is important because meaning itself is not a static or stable monolith. It circulates through society and changes over time. Meaning is intrinsically on the move and can be understood as effect of signs and symbols and is tied to cognitive and environmental motivation. Cognitive motivation includes the aspects of culture an individual has internalized. Environmental motivation likely motivates meaning as well. Therefore, motivation arises from the ongoing interaction between the individual and the environment. [@Norton2018] Dr. Norton creates a useful pragmatist view of the meaning identifying three categories; "What matters?", "What are you going to do about it?", and "Why?" As members of society we draw from our experiences through intersectionality to answer these questions. It is additionally interesting when considering the social meaning of something as old in human history as religious and spiritual meaning due to the temporal aspect of meaning. 

Allowing meaning as a concept to shift to an occurrence in the mind and/or environment allows meaning to become a processual category and it suggests meaning manifests as actions, events, ideas, habits, dispositions, and even more collections of signs and social regularities. [@Norton2018] The trends in religious and spiritual beliefs may be tied to this phenomenon of meaning on the move and its interactions between individuals and their environment. As the social and educational landscapes of the United States changed over the last century, so did the social meaning of religious/spiritual beliefs and practices. These changes may be measured through analyzing associations between trends in religiosity, educational factors, and social factors over time. Existing trends in previous research provides interesting clues and questions.

A study examining GSS data from 1972-2016, found interesting trends in the beliefs and practices of the religiously affiliated and the unaffiliated in the United States. From 1973 to 2000, belief in the afterlife increased for both the religiously affiliated and the unaffiliated. However, a clear divergence emerges in the year 2000. The belief in the afterlife for the affiliated increased while it decreased for the unaffiliated. This same trend was found in the belief in God. [@Gullickson2018] Dr. Gullickson explains the population of the unaffiliated is composed of those who were never affiliated and those who are disaffiliated and adds after the year 2000, the population of the never affiliated seemed to level off in terms of growth. However, the population of the disaffiliated grew substantially. This divergence is interesting. Is this shift associated with the social meaning of religious/spiritual beliefs and practices on the move as interacting between individuals and their environment? If so how and what factors of intersectionality are potentially involved? Care must be utilized when measuring these associations. Correlation does not mean causation. However, understanding more about how the intersecting aspects of human social identity influences how meaning moves may provide clues on how other areas of social construction changes over time as well. 

In a study involving 626 participants, the influence of education, sex, and age on religiosity was conducted. The results were examined across three cohorts, the Baby Boomer generation, Generation X, and Generation Y. Results showed females in the Baby Boomer and Generation X scored higher on the intrinsic religiosity than males. A positive relation was found between education and both daily spiritual experiences and intrinsic religiosity with a negative relation found between education and extrinsic religiosity. Additionally, Generation Y reported higher levels of extrinsic religiosity and Generation Y females socred lower on daily spiritual experiences and intrinsic religiosity than males. [@Green2013] This is interesting because this observed pattern seems similar to the divergence noted between the religiously affiliated and unaffiliated in Dr. Gullickon's research. Education and gender are not the only intersecting factor either.

AN international study of religiosity and education found interesting associations with GDP. When considering the process of modernization, researchers postulated the objective of modernization is economic growth and is adjacent to a secularization process. However, when examining the relation between education and religiosity it was found education served as a mediator between religiosity and GDP. This may account for some of the variation found between countries as some have religion more heavily integrated into education. [@Ha2019] The level of integration between religion and education would likely produce different results. If the social meaning of religiosity itself is the result of cognitive and environmental motivaiton we shold expect to see differing results from the influence of modernization on religious/spiritual beliefs and practices. Individuals in different social environments will have different motivations when utilizing the three categories of meaning. This is the importance of intersectionality in constructing social meaning while also allowing it to move through both groups and time.  

A longitudinal analysis of the effects of higher education on religiosity found a decline in religiosity upon achieving a Bachelors degree. The same study also found an increase in religiosity scores among respondents as they aged. through adulthood. This implies two things. First, people may become less religious as they enter college and leave direct parental influence. This may be influenced by social and academic curriculum. Second, this implies after earning a Bachelors degree and moving tthrough adult life, the process of starting a family may influence religious patterns among aging adults in a positive relationship. [@Schwadel2016] However, it should also be noted in another study examining the cross-national effects of higher education on religiosity, it was found GDP served as a mediator between national-level education and national-level religiosity[@Shwadel2015] This suggests age has a distinctive influence on religiosity as well as the social and academic environment of college. Once again different aspects of human social identity potentially intersect to facilitate the convergence between cognitive and environmental motivations allowing the social meaning of religious/spiritual beliefs and practices to change over time. 

The social meaning of religious/spiritual beliefs and practices are influenced by intersectionality because they are areas of convergence between cognitive and environmental motivations. This allows the meaning to change over time as both a person and a society change over time. The United States provides an interesting population to study the potential evolution of the social meaning of religious and spiritual importance due to the diverse religious affiliations, large-scale educational access, and diverse cultural backgrounds. The question of interest to this research is, how does the intersections of educational attainment and racial identification potentially influence religiosity?
Given the consideration of the previous areas of research mentioned, three main hypothesis emerge. First, as educational attainment increases, religiosity will decrease. If there is an environmental motivation factor influencing religiosity and it is observed to decrease upon entering college, religiosity should decrease as education continues. Second, because all groups in the United States do not experience the same access to the large-scale educational system, differences should be seen in religiosity among educational attainment across racial categories. Third, because these factors intersect with the specific religion with which a respondent is affiliated, (or even in the case where that religion is none) differences in religiosity scores should be seen across educational attainment and racial identification when controlling for religious affiliation. 

# Data and Methods
This research project utilizes a repeated cross-sectional study design focusing on GSS (General Social Survey) data from 2006 to 2018. These years were chosen for the availability of data concerning questions used to construct a measure of religiosity. The questions used to construct a measure of religiosity are frequency of service attendance, frequency of prayer, respondent's belief in God, and how important is religion to the individual respondent. Service attendance is recorded as an ordinal variable with the following categories; never, yearly, monthly, weekly, and more than weekly. Frequency of prayer is recorded as an ordinal variable with the following categories; never, weekly, daily, more than daily. The respondent's belief in God is recorded as an ordinal variable with the following categories; no belief, belief in a higher power, belief but doubts, and believes with no doubt. The measure of religious importance to the individual respondent is recorded as an ordinal variable with the following categories; no religion, not strong, somewhat strong, and strong. These variables were used in a summated scale to create the overall measure of religiosity. 

Educational attainment among respondents was collapsed into the following categories; less than high school, high school, some college (including trade certifications and associate degrees), bachelors degree, and graduate degree (including masters degrees and PhD's). The racial identification of respondents were collapsed into the following categories; white, black, latinx, asian, and asian/pacific islander. In order to control for religious affiliation, the following categories of affiliation were considered; Protestant, Catholic, Other Christian (which includes the categories "Christian", "Orthodox Christian", and "Inter-nondenominational"), Jewish, Muslim, Buddhist, Hindu, other, (which includes Indigenous and other Eastern religions) and none. In order to consider the temporal intersection, age was both kept as a quantitative variable and collapsed in ten-year-age groups with the exception of the first age group, 18-29. This group was constructed due to a small amount of respondents in the 18-19 year-old age group. Respondent's gender was recorded as male or female as that is the format of the GSS. Cases with missing data were dropped to allow for a complete-case analysis of the remaining observations. All models were created as general linear models. However, the age squared term in models one through four control for the non-linear effects of age. 

# Limitations
There are several limitations to consider in this research. First, correlation is not causation. Even potentially finding relationships between religiosity, education, and race does not mean education and/or race "cause" an effect. There can be a measure of association, but likely other factors contribute as well. Due to the nature of intersectionality, it is likely numerous intersections play a role. This research project is only limited to examining the potential association between the identified variables. Second, race itself is not a physical category, it is a social construction. This means any influence or association is not the result of some biological factor and more likley associated with cultural factors for both the individual group and the influence of the larger collective. Finally, the very measure of religiosity may be flawed in that it does not consistently measure the individual variables used to create religiosity across all religions. This may lead to systematic under of over counting with consideration of these intersections. 

# Potential Issues in Measuring Religiosity
There are two specific areas of interest when identifying the limitations of this study; religious attendance and the belief in God of each respondent. This must be aknowleged because these are two of the four categories used to construct the religiosity measure. More so, they are ordinal variables and therefore have a value rating in a specific order. There is no set specific measure of religiosity. However, the four selected for this research are four of the most commonly used categories.  

Measuring religious attendance was done by asking respondents how often they attended religious service. The given options to respond were; never, yearly, monthly, weekly, and more than weekly. This means a respondent attending service at a weekly rate will score higher in the service attendance category than a respondent who indicates they attend monthly. This works fine as long as each religion and each intersection is measured equaly. However, each religion and each culture has slight differences in practices and beliefs.

```{r}
#| label: fig-Religious_Attendance_by_Religious_Affiliation
#| fig-cap: Religious Attendance by Religious Affiliation
gss |>  
  ggplot(aes(x=relig_attend, y=after_stat(prop), group=relig_affil, fill=relig_affil))+
  geom_bar(position="dodge")+
  #facet_wrap(~race)+
  #scale_fill_viridis_d()+
  scale_fill_manual(values=c("#154733","#FEE123","red", "blue", "black", "orange", "grey", "pink", "purple" ))+
  scale_y_continuous(labels=scales::percent)+
  labs(x=NULL, y=NULL, 
       title="Religious Attendance by Religious Affiliation")+
  coord_flip()+
  theme_bw()
```

As seen in @fig-Religious_Attendance_by_Religious_Affiliation there are two categories highly represented as never attending service; "other", and "Buddhist". This can be problematic as service requirements and structure may not be the same for these two groups as they are for the other represented affiliations. Because this is an ordinal variable contributing it's score to the overall religiosity measure, these two groups may be systematically under-considered. The "other" category includes indigenous and other eastern religions. It could be problematic to value their service attendance rating in this manner. If a religion does not have within it's structure a need to go to a specific service at a specific time, it potentially underscores the actual religiosity of those in that group. It also prioritizes the Christian religious format because service attendance is, on average, a large part of that religious practice. This is not the only potential issue with using religious attendance as a variable to create a measure of religiosity. There are differences across gender as well.

```{r}
#| label: fig-Religious_Attendance_by_Respondents_Sex_and_Religious_Affilation
#| fig-cap: Religious Attendence by Respondents Sex and Religious Affiliation
gss |>  
  ggplot(aes(x=relig_attend, y=after_stat(prop), group=resp_sex, fill=resp_sex))+
  geom_bar(position="dodge")+
  facet_wrap(~relig_affil)+
  #scale_fill_viridis_d()+
  scale_fill_manual(values=c("#154733","#FEE123"))+
  scale_y_continuous(labels=scales::percent)+
  labs(x=NULL, y=NULL, 
       title="Religious Attendance by Religious Affiliation and Respondent's Sex")+
  coord_flip()+
  theme_bw()
```

As shown in @fig-Religious_Attendance_by_Respondents_Sex_and_Religious_Affilation it is seen men and women do not all have the same service attendance habits across religions. For example, within the Muslim faith it can be seen men attend service more often than women. This is most noticeable in the "weekly" category for attendance. Within this category, men are over represented compared to women. There are likely cultural explanations for this difference in service attendance. This difference in attendance needs to be considered when using attendance as a composite in the religiosity variable. Additionally, The Buddhist faith seems to show a similar pattern. Women are over represented in the "never" category while men are the majority in each other category. Again, there is likely a systematic misrepresentation of attendance based on gender factors. 

In the case of religious attendance used as a composite measure of religiosity there may be concern for misrepresentation in religious importance. Differences in attendance resulting from cultural structures or the service structure of particular religions do not necessarily mean one group is more or less religious than another. It simply means there are cultural or structural differences in how a religion is integrated into a particular society. This is not the only potential issue. Belief in god could also be considered in this way.

```{r}
#| label: fig-Beleif_in_God_by_Religious_Affiliation
#| fig-cap: Belief in God of Respondents by Religious Afiliation.
gss |>  
  ggplot(aes(x=believe_god, y=after_stat(prop), group=relig_affil, fill=relig_affil))+
  geom_bar(position="dodge")+
  #facet_wrap(~race)+
  #scale_fill_viridis_d()+
  scale_fill_manual(values=c("#154733","#FEE123","red", "blue", "black", "orange", "grey", "pink", "purple" ))+
  scale_y_continuous(labels=scales::percent)+
  labs(x=NULL, y=NULL, 
       title="Belief in god by Religious Affiliation")+
  coord_flip()+
  theme_bw()
```

As shown in @fig-Beleif_in_God_by_Religious_Affiliation it is seen Buddhists are the majority representation in the belief in a higher power. However, this could be an issue because it implies a belief in a higher power is less religious than a belief in a god. This is an ordinal variable giving more value to the answers professing a belief in a specific god. as with service attendance, the "other" category must also be considered. It may be problematic to place indigenous beliefs below Christian because they do not incorporate the same or a single deity belief system. That, in and of itself, is not enough to classify them as less religious. 

In each of these cases it is problematic to use the current, and most commonly used, measures of religiosity. They seem to prioritize Christian religious beliefs and practices. This may not give us a measure of religiosity usable across religious affiliations or gender. This project utilized 4 of the 5 most common variables used to create a measure of religiosity. If two variables potentially have issues with under or over representing certain groups due to cultural or religious structural reasons, the overall measure should contain a rather large margin of error preventing accurate predictions of associations. It may be best to consider options allowing the measure of religiosity to be done within religions instead of across them. This would require internal considerations for each affiliation to account for cultural and structural differences, then using internal measures to create an overall religiosity applicable to specific religious/spiritual belief systems. This could potentially provide a more robust analysis of aggregate religiosity. Meaning is on the move and it is seen in these examples. As the social meaning of religious/spiritual beliefs and practices moves through society it is incorporated differently depending on the intersections of human social identity. We can potentially get a better measure and understanding of this process in the context of religious structures by comparing the accurate differences and similarities between them. But, when considering the value or importance of a respondent's religion, we need to consider the differences more carefully and work to place the right context on them.  

# Results
The focus of this research is to examine the possible associations between religiosity, education, and racial identification. There are three main hypothesis postulated. First, as educational attainment increases, religiosity will decrease. Second, because all groups in the United States do not experience the same access to the large-scale educational system, differences should be seen in religiosity among educational attainment across racial categories. Third, because these factors intersect with the specific religion with which a respondent is affiliated, differences in religiosity scores should be seen across educational attainment and racial identification when controlling for religious affiliation. These differences may reflect the process of meaning on the move. As each group in the United States integrates their social identity they go through the process of integrating the three categories of the pragmatist view of meaning; "What matters?", "What are you going to do about it?", and "Why?" During this process, each intersection of human social identity comes into play creating the foundation for both internalizing the social meaning of religious/spiritual beliefs and practices. Here, both the cognitive and environmental motivations are potentially observed. The process of internalizing the social meaning of religious/spiritual beliefs and practices happens at a cognitive level. The intersectional aspect of human social identity provides a social environmental consideration. This does not mean physical environments do not have an influence. Rather, the focus on this research is to observe trends potentially associated with the intersectionality contribution as it relates to education and racial identification and it's influence on religiosity. It is good to begin looking at religiosity across the distribution of educational attainment. 

```{r}
#| Label: fig-Religiousity_by_Educational_Attainment
#| fig-cap: Religiousity by Repsondents Educational Attainment
gss |>
  group_by(educ_attain) |>
  summarize(religiosity=mean(religiosity)) |>
  ungroup() |>
  ggplot(aes(x=educ_attain, y=religiosity, fill=educ_attain))+
  geom_col(position="dodge")+
  #scale_fill_viridis_d()+
  scale_fill_manual(values=c("#154733","#FEE123","red", "blue", "black", "orange", "grey", "pink", "purple"))+
  labs(x=NULL, y=NULL, 
       title="Religiosity by Educational Attainment")+
  theme_bw()
```

When examining the relation between the religiosity and educational attainment among respondents, the results appear to support hypothesis one. As seen in @fig-Religiousity_by_Educational_Attainment religiosity scores decrease as educational attainment increases. This is of interest because it may suggest a social environment in education facilitating the negative relationship with religiosity. However, caution is needed as correlation is not causation. There are many factors to consider within educational attainment as previously mentioned. Each group in the United States has varying access to different levels of educational attainment. An examination of eduacational attainment across racial categories will be helpful in observing trends within the context of intersectionality. 

```{r}
#| label: fig-Educational_Attainment_by_Race
#| fig-cap: Level of Educational Attainment of Respondents by Racial Identification.
gss |>  
  ggplot(aes(x=educ_attain, y=after_stat(prop), group=race, fill=race))+
  geom_bar(position="dodge")+
  #facet_wrap(~race)+
  #scale_fill_viridis_d()+
  scale_fill_manual(values=c("#154733","#FEE123","red", "blue", "black" ))+
  scale_y_continuous(labels=scales::percent)+
  labs(x=NULL, y=NULL, 
       title="Education Level by Racial Identification")+
  coord_flip()+
  theme_bw()
```

There are some trends to note when examining the association between educational attainment and racial identification.As shown in @fig-Educational_Attainment_by_Race educational attainment varies across racial identification. The Asian and White racial categories are among the most represented in the educational categories of bachelors and graduate degrees. The Black and Asian/Pacific Islander categories represent the majority withing he some college categories. The Black and White represent the majority in the high school category. Finally, the latinX group represents the majority in the less than high school category. There are clear differences in educational attainment across racial identifications. This is helpful as it relates to the foundation of hypothesis two. Different racial groups have differing access to educational attainment. This should produce differences in religiosity among educational attainment across racial categories. In terms of intersectionality, a respondents religiosity should be influenced by the intersection of their racial identification and specific educational attainment. 

```{r}
#| Label: fig-Religiousity_by_Education_and_Race
#| fig-cap: Religiousity by Repsondents Education Level and Race
gss |>
  group_by(race, educ_attain) |>
  summarize(religiosity=mean(religiosity)) |>
  ungroup() |>
  ggplot(aes(x=race, y=religiosity, fill=educ_attain))+
  geom_col(position="dodge")+
  #scale_fill_viridis_d()+
  scale_fill_manual(values=c("#154733","#FEE123","red", "blue", "black", "orange", "grey", "pink", "purple"))+
  scale_y_continuous(labels=scales::percent)+
  labs(x=NULL, y=NULL, 
       title="Religiosity by Educational Attainment and Race")+
  theme_bw()
```

Trends begin to emerge when examining religiosity across educational attainment and racial identification. As shown in @fig-Religiosity_by_Education_and_Race there are clear differences in religiosity across both factors. This seems to support hypothesis two. Among the White racial group, religiosity appears to decrease as educational attainment increases. A similar pattern seems to emerge for the Asian racial category. The LatinX racial group seems is of interest because it appears religiosity in that group may not be influenced by the intersection of educational attainment and racial identification as much as the other groups. The Black racial category appears to have the opposite relation than that of Whites and Asians. As educational attainment increases religiosity also increases. This is an interesting results. To better understand some of the trend among Black respondents in this context Table_of_Religiosity_by_Education_and_Race it appears the trend may be driven by those not affiliated with a religious group become more religious as they attain education, on average. Graphs and tables help to understand some trends. However, in this context, they are more helpful in defining the questions related to the hypothesis presented. To better understand, models are needed.  


```{r}
#| label: table-Table_of_Religiosity_by_Education_and_Race
#| fig-cap: Religiosity_table_by_education_and_race
round(prop.table(table(gss$relig_affil, gss$educ_attain, gss$race),c(2,3)),3)*100
```

# Religiosity Models
  
```{r}
#| label: Stair-step approach for models
ordered_factor <- function(fact_var) {
  ord_fact <- factor(fact_var, ordered=TRUE)
  categories <- levels(fact_var)
  n_cat <- length(categories)
  cont <- matrix(0, n_cat, n_cat-1)
  cont[col(cont)<row(cont)] <- 1
  rownames(cont) <- categories
  colnames(cont) <- paste(categories[2:n_cat], categories[1:(n_cat-1)],
                          sep=" vs. ")
  contrasts(ord_fact) <- cont
  return(ord_fact)
}

```

```{r}
#| label: Stair-step approach for previous table
#| fig-cap: Religiosity_stair_step_table_by_education_and_race
gss$educ_attain <- ordered_factor(gss$educ_attain)
contrasts(gss$educ_attain)
```
  
```{r}
#| Label: Modeling Religiosity 
model1 <- glm(religiosity~educ_attain*age+I(age^2), data=gss)
model2 <- update(model1, .~.+educ_attain*race, data=gss)
model3 <- update(model2, .~.+resp_sex*race, data=gss)
model4 <- update(model3, .~.+relig_affil)
```

```{r}
#| label: fig-Rendering_Models
#| fig-cap: Models of religiosity by education and Race with interaction terms.
#| results: asis
#to print to screen replace texreg with screenreg
texreg(list(model1, model2, model3, model4),
        custom.coef.names = c("Intercept", "high school vs less than high school", "some college vs high school",
        "bachelors vs some college", "graduate vs bachelors", "age", "age squared", "high school vs less than high school controlling age",
        "some college vs high school controlling age", "bachelors vs some college controlling age", "graduate vs bachelors controlling age",
        "black", "latinX", "asian", "asian/pi", "high school vs less than high school among Black", "some college vs high school among Black",
        "bachelors vs some college among Black", "graduate vs bachelors among Black", "high school vs less than high school among LatinX",
        "some college vs high school among LatinX", "bachelors vs some college among LatinX", "graduate vs bachelors among LatinX",
        "high school vs less than high school among Asian", "some college vs high school among Asian", "bachelors vs some college among Asian",
        "graduate vs bachelors among Asian", "high school vs less than high school among Asian/PI", "some college vs high school among Asian/PI",
        "bachelors vs some college among Asian/PI", "graduate vs bachelors among Asian/PI", "male", "Black Male", "LatinX Male",
        "Asian Male", "Asian/PI Male", "Catholic", "Other Christian", "Jewish", "Muslim", "Buddhist", "Hindu", "Other", "None"),
       digits = 3,
       caption.above=TRUE, 
       include.rsquared=TRUE,
       include.adjrs=FALSE,
       include.nobs=TRUE,
       include.rmse=FALSE)
```

As seen in @fig-Rendering_Models Model One shows a negative relationship between religiosity and educational attainment. However, when controlling for age, the intensity of this association drops to almost zero. This challenges hypothesis one. As educational attainment rises from less than high school through some college, a positive relation seems to emerge. Then, upon attaining a graduate degree, religiosity scores drop, on average. There is no clear trend of religiosity declining as education increases. This can be consistent with the research by Schwadell focusing on educational and age influences. If religiosity increases with age, on average, we could expect to see a leveling effect on the educational influence. 

Model two additionally includes educational attainment across racial categories. In this model, hypothesis one is not supported. There is a positive overall relations between religiosity and educational attainment, on average. This is again dropped to near zero when controlling for age. When controlling for racial categories, there are differences in religiosity observed as well. When compared to the White racial category, the same pattern is observed in the Black racial category as seen in the previous graph. There is a positive relation between religiosity and educational attainment. This suggests as Black people attain higher education, religious/spiritual beliefs and practices become more important, on average. When examining the LatinX category, there is a negative relation between religiosity and education except when moving from some college and attaining a bachelors degree. Model two suggests a positive relation in religiosity upon attaining a bachelors degree. Model two also shows the Asian category to have a similar overall pattern to the LatinX category. A general negative relation between religiosity and education was observed except when respondents entered college. That educational shift produced a positive relation between religiosity and education. Model two provides support for hypothesis two. There are differences in religiosity across educational and racial intersections. Model three adds gender as a consideration to try and understand some of the influence discussed within the limitations. 

Model four looks at the heart of hypothesis three. Differences in religiosity across educational attainment and racial categories should vary when controlling for religious affiliation. Model four begins with an overall positive relation between religiosity and education, not supportive of hypothesis three. When controlling for the effects of age, this difference begins to drop away. When compared to the White racial category, controlling for age and racial category, respondents in the Black and LatinX were found to have significantly higher religiosity scores, on average. Model four also shows, when compared to Protestants, religiosity trends were not consistent across all religious affiliations. Compared to Protestants, while controlling for age, racial categories, and educational attainment, the Other Christian category and the Muslim category showed positive religiosity scores, on average. This supports hypothesis three as there are observed differences in religiosity among respondents across educational attainment and racial categories when contorlling for religious affiliation.   

# Conclusions
 Hypothesis one was not supported by the modeling outcomes. There was no clear decreasing trend in religiosity as educational attainment increased. In many instances, religiosity seemed to increase as educational attainment increased. Hypothesis two and three were supported by the modeling outcomes as they showed clear differences in religiosity across educational outcomes when controlling for racial categories. Black respondents showed increased religiosity as educational attainment increased compared to Whites, on average. There were differences in religiosity scores when controlling for religious affiliation as well. Other Christian and Muslim religious affiliations showed a positive association between religiosity and educational attainment, while controlling for racial identification, on average. This support hypothesis three. This indicates the potential influence intersectionality has on religiosity.
 
The social meaning of religious/spiritual beliefs and practices is on the move through groups and over time. This is because of the cognitive and environmental motivations involved in integrating these beliefs and practices into an individual and collective human structure. These groups approach the three categories of he pragmatist view of meaning, "What matter?", "What are you going to do about it?", and "Why?" [@Norton2018] not just as a respondent, but as male or female of a particular racial category, attaining a specific level of education, belonging to a certain religion. These intersections provide the social environmental motivation. The differences in religiosity scores observed across education and racial identification, while controlling for religious affiliations, show the process of the social meaning of religious/spiritual practices evolving and changing as it moves across groups and time. This is important to note because this process happens with every social construct made. Understanding why and how these meanings change gives insight into how needed changes can occur. This may provide insight into how changes in the social meaning of race and gender may occur. 
